---
title: "Quickstart for R and the DIDE cluster"
author: "Rich FitzJohn"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Quickstart for R and the DIDE cluster}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

> Get yourself running R jobs on the cluster in 10 minutes or so.

Assumptions that I make here:

* you are using R

* your task can be represented as running a function on some inputs
  to create an output (a file based output is OK)

* you are working on a network share and have this mounted on your
  computer

* you know what packages your code depends on

* your package dependencies are all on CRAN, and are all available
  in windows binary form.

If any of these do not apply to you, you'll probably need to read
the full vignette.  In any case the full vignette contains a bunch
more information anyway.

## Install a lot of packages

On windows in particular, this step can be infuriating if something
triggers an upgrade in a package that is being depended on.  If you
end up in a situation where packages simply can't be loaded you may
need to restart R and try installing things from scratch.
Hopefully I can come up with a way of reducing the pain here.

``` {r eval=FALSE}
install.packages("devtools")
devtools::install_github(c(
  "richfitz/ids",
  "dide-tools/context",
  "richfitz/queuer",
  "dide-tools/didewin"))
```

Or:
``` {r eval=FALSE}
drat:::add("richfitz")
install.packages("didewin")
```

Or:
``` {r eval=FALSE}
install.packages("didewin", repos=c(CRAN="https://cran.rstudio.com",
                                    drat="https://richfitz.github.io/drat"))
```

## Describe your computer so we can find things

On windows if you are using a domain machine, you should need only
to select the cluster you want to use

``` {r eval=FALSE}
didewin::didewin_config_global(cluster="fi--didemrchnb")
```

Otherwise, and on any other platform you'll need to provide your username:
``` {r }
didewin::didewin_config_global(username="rfitzjoh",
                               cluster="fi--didemrchnb")
```

If this is the first time you have run this package, best to try
out the login proceedure with:
``` {r }
didewin::web_login()
```

## Describe your project dependencies so we can recreate that on the cluster

Make a vector of packages that you use in your project:
``` {r }
packages <- c("ape", "MASS")
```

And of files that define functions that you ned to run things:
``` {r }
sources <- c("mysources.R", "utils.R")
```

Then save this together to form a "context".
``` {r }
ctx <- context::context_save("contexts", packages=packages, sources=sources)
```

## Build a queue, based on this context.

This will prompt you for your password, as it will try and log in.
``` {r }
obj <- didewin::queue_didewin(ctx)
```

Once you get to this point we're ready to start running things on
the cluster.  Let's fire off a test to make sure that everything works OK:
``` {r }
t <- obj$enqueue(sessionInfo())
```

We can poll the job for a while, which will print a progress bar.
If the job is returned in time, it will return the result of
running the function.  Otherwise it will throw an error.
``` {r }
t$wait(120)
```

You can use `t$result()` to get the result straight away (throwing
an error if it is not ready) or `t$wait(Inf)` to wait forever.
