---
title: "Quickstart for R and the DIDE cluster"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Quickstart for R and the DIDE cluster}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

<!-- HEADER -->

```{r include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  error = FALSE,
  comment = "#>"
)
r_output <- function(x) {
  cat(c("```r", x, "```"), sep = "\n")
}
```

> Get yourself running R jobs on the cluster in 10 minutes or so.

Assumptions that we make here:

* you are using R

* your task can be represented as running a function on some inputs
  to create an output (a file based output is OK)

* you are working on a network share and have this mounted on your
  computer

* you know what packages your code depends on

* your package dependencies are all on CRAN, and are all available
  in windows binary form.

If any of these do not apply to you, you'll probably need to read
the full vignette.  In any case the full vignette contains a bunch
more information anyway.

## Install a lot of packages

Install the packages using [`drat`](https://cran.rstudio.com/package=drat)

```r
# install.package("drat") # if you don't have it already
drat:::add("mrc-ide")
install.packages("didehpc")
```

## Describe your computer so we can find things

On windows if you are using a domain machine, you should need only
to select the cluster you want to use

```r
options(didehpc.cluster = "fi--didemrchnb")
```

Otherwise, and on any other platform you'll need to provide your username:

```r
options(didehpc.cluster = "fi--didemrchnb",
        didehpc.username = "yourusername")
```

You can see the default configuration with

```{r}
didehpc::didehpc_config()
```

If this is the first time you have run this package, best to try
out the login procedure with:

```{r}
didehpc::web_login()
```

because this exposes a number of problems early on.

## Describe your project dependencies so we can recreate that on the cluster

Make a vector of packages that you use in your project:

```{r}
packages <- c("dplyr", "tidyr")
```

And of files that define functions that you ned to run things:

```{r}
sources <- "mysources.R"
```

If you had a vector here that would be OK too.

Then save this together to form a "context".

```{r}
ctx <- context::context_save("contexts", packages = packages, sources = sources)
```

If you have no packages or no sources, use `NULL` or omit them in
the call below (which is the default anyway).

The first argument here, `"contexts"` is the name of a directory
that we will use to hold a lot of information about your jobs.  You
don't need (or particularly want) to know what is in here.

## Build a queue, based on this context.

This will prompt you for your password, as it will try and log in.

It also installs windows versions of all packages within the
`contexts` directory -- both packages required to get this whole
system working and then the packages required for your particular
jobs.


```{r}
obj <- didehpc::queue_didehpc(ctx)
```

Once you get to this point we're ready to start running things on
the cluster.  Let's fire off a test to make sure that everything works OK:

```{r}
t <- obj$enqueue(sessionInfo())
```

We can poll the job for a while, which will print a progress bar.
If the job is returned in time, it will return the result of
running the function.  Otherwise it will throw an error.

```{r}
t$wait(120)
```

You can use `t$result()` to get the result straight away (throwing
an error if it is not ready) or `t$wait(Inf)` to wait forever.

```{r}
t$result()
```

## Running a single task

This is just using the `enqueue` function as above.  But it also
works with functions defined in files passed in as `sources`; here
the function `random_walk`.

```{r}
t <- obj$enqueue(random_walk(0, 10))
res <- t$wait(120)
res
```

The `t` object has a number of other methods you can use:

```{r}
t
```

Get the result from running a task

```{r}
t$result()
```

Get the status of the task

```{r}
t$status()
```

(might also be "PENDING", "RUNNING" or "ERROR"

Get the original expression:

```{r}
t$expr()
```

Find out how long everything took

```{r}
t$times()
```

You may see negative numbers for "waiting" as the submitted time is
based on your computer and started/finished are based on the
cluster.

And get the log from running the task

```{r}
t$log()
```

There is also a bit of DIDE specific logging that happens before
this point; if the job fails inexplicably the answer may be in:

```{r}
obj$dide_log(t)
```

Want more information? See `vignette("didehpc")` for more details.
